---
title: "Preparing Agriculture Data for Maxent"
author: "Nakoa Farrant"
date: "11/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following method is based on: https://www.nceas.ucsb.edu/scicomp/usecases/PolygonDissolveOperationsR

This script should demonstrate how to dissolve polygons using maptools and do area calculations using PBSmapping

## 1. First include the relevant libraries:

```{r, message=FALSE}
library(maptools) # for geospatial services; also loads foreign and sp packages
library(gpclib) # General polygon clipping library
library(rgdal) # for map projection work; also loads sp
library(PBSmapping) # for GIS-like geospatial object manipulation and analysis including poly
library(sp)
library(tidyverse)
library(ggspatial)
```

The shapefile folder was included in the hawaii-agriculture GitHub repository as a zip file. Once the repo is forked, one should be able to unzip the file which will produce a folder containing the shapefile and all of its dependencies. 

The 2015 Agricultural Baseline shapefile data that is used here was downloaded from the Hawai‘i Office of Planning's website that is compiled for the Hawai'i Statewide GIS Program: http://planning.hawaii.gov/gis/download-gis-data/. 

This agricultural baseline was generated by Ryan L. Perroy, Jeffrey Melrose, and Sylvana Cares for the Hawai'i Dept of Agriculture. A report on this data can be found in The evolving agricultural landscape of post-plantation Hawai‘i published in Applied Geography in Nov 2016


## 2. Read in the shapefile data
```{r}
hi_ag_base <- readOGR("2015AgBaseline.shp") # Large SpatialPolygonsDataFrame
```

## 3. Project the polygons on appropriate UTM projection
Transform the polygons that were read in as unprojected geographic coordinates to a UTM projection
```{r}
hi_ag_proj <- spTransform(hi_ag_base, CRS("+proj=utm +ellps=GRS80 +datum=NAD83"))
```

## 4. Subset the field shapes by crop category

This isolates the fields based on the crop type that they were labeled with

```{r}
aqua_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Aquaculture")
banana_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Banana")
coffee_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Coffee")
dairy_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Dairy")
divcrop_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Diversified Crop")
flowers_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Flowers / Foliage / Landscape")
forestry_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Commercial Forestry")
macnuts_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Macadamia Nuts")
papaya_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Papaya")
pasture_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Pasture")
pineapple_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Pineapple")
seeds_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Seed Production")
sugar_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Sugar")
taro_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Taro")
fruits_poly = subset(hi_ag_proj, hi_ag_proj@data$CropCatego == "Tropical Fruits")

```

## 5. Generate random presence points for each crop

The maximum entropy model that will be used to analyze future crop distributions requires random presence points where fields containing each crop are currently located. 

```{r}
# Should the number of random presence points selected vary with the total area of each crop type? 

aqua_pres = spsample(aqua_poly, n = 1e2, "random")
banana_pres = spsample(banana_poly, n = 1e2, "random")
coffee_pres = spsample(coffee_poly, n = 1e2, "random")
dairy_pres = spsample(dairy_poly, n = 1e2, "random")
divcrop_pres = spsample(divcrop_poly, n = 1e2, "random")
flowers_pres = spsample(flowers_poly, n = 1e2, "random")
forestry_pres = spsample(forestry_poly, n = 1e2, "random")
macnuts_pres = spsample(macnuts_poly, n = 1e2, "random")
papaya_pres = spsample(papaya_poly, n = 1e2, "random")
pasture_pres = spsample(pasture_poly, n = 1e2, "random")
pineapple_pres = spsample(pineapple_poly, n = 1e2, "random")
seeds_pres = spsample(seeds_poly, n = 1e2, "random")
sugar_pres = spsample(sugar_poly, n = 1e2, "random")


```

Plot the spatial polygons for each rop type and also plot the random points that were generated for each 


```{r}
plot(aqua_poly)
points(aqua_pres, pch = 3)
title("Aquaculture")

plot(banana_poly)
points(banana_pres, pch = 3)
title("Banana")

plot(coffee_poly)
points(coffee_pres, pch = 3)
title("Coffee")

plot(dairy_poly)
points(dairy_pres, pch = 3)
title("Dairy")

plot(divcrop_poly)
points(divcrop_pres, pch = 3)
title("Diversified Crops")

plot(flowers_poly)
points(flowers_pres, pch = 3)
title("Flowers")

# The code hangs on the forestry plot for some reason but runs at reasonable speed for all other crop types
plot(forestry_poly)
points(forestry_pres, pch = 3)
title("Commercial Forestry")

plot(macnuts_poly)
points(macnuts_pres, pch = 3)
title("Macadamia Nuts")

plot(papaya_poly)
points(papaya_pres, pch = 3)
title("Papaya")

plot(pasture_poly)
points(pasture_pres, pch = 3)
title("Pasture")

plot(pineapple_poly)
points(pineapple_pres, pch = 3)
title("Pineapple")

plot(seeds_poly)
points(seeds_pres, pch = 3)
title("Seeds")

plot(sugar_poly)
points(sugar_pres, pch = 3)
title("Sugar")

```

## 6. Create shapes for the background (absence) points

## 7. Condition background for each crop type

Use information on suitable temperature, elevation, soil type, precipitation, and proximity to streams or other water sources to determine the suitability of different swaths of land

There may not be any data on spring locations which could provide a sufficient water supply in otherwise unsuitable land for crops that have higher water demand

